Body size patterns

### 2019-06-27 MS_Analysis_3spp_4rivers_Patterns_from 2spp_bySpp.Rmd
update model selection for gams

```{r Load data, message=FALSE}
library(itsadug)
library(arm)
library(relaimpo)
library(linkedModels)
library(arrayhelpers)
library(mgcv)
library(tidyverse)
library(gridExtra)

@@

# 2-way
vB$speciesRiverOrdered <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
vB$speciesCohortOrdered <- factor(paste(vB$speciesOrdered,vB$cohortOrdered, sep = '_'), ordered = T)
vB$cohortRiverOrdered <- factor(paste(vB$cohortOrdered,vB$riverOrdered, sep = '_'), ordered = T)

@@

vB$salmon01 <- ifelse(vB$cohort <=2004, 1, 0)

@@

ggplot(cohortStrength, aes(cohort, cohortStrengthScaled, group = species, linetype = species, shape = species)) +
  geom_point(size = 2) +
  geom_line(size = 1.1, color = 'grey40') +
  labs(x = 'Cohort',
       y = 'Cohort strength') +
  theme_publication() +
  facet_grid(~metaPop01)

@@

```{r gam models - only need to run for model selection}
####################################
# gam models

# base model

mod <- bam(observedLength ~ cohort * river + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")
summary(mod)

>>>>

```{r Load data, message=FALSE}
library(itsadug)
library(arm)
library(relaimpo)
library(GGally)
library(linkedModels)
library(arrayhelpers)
library(mgcv)
library(tidyverse)
library(gridExtra)

@@

# 2-way
vB$speciesRiverFactor <- factor(paste(vB$speciesFactor,vB$riverFactor, sep = '_'))
vB$cohortRiverFactor <- factor(paste(vB$cohortFactor,vB$riverFactor, sep = '_'))
vB$seasonRiverFactor <- factor(paste(vB$season,vB$riverFactor, sep = '_'))

vB$speciesRiverOrdered <- factor(paste(vB$speciesOrdered,vB$riverOrdered, sep = '_'), ordered = T)
vB$speciesCohortOrdered <- factor(paste(vB$speciesOrdered,vB$cohortOrdered, sep = '_'), ordered = T)
vB$cohortRiverOrdered <- factor(paste(vB$cohortOrdered,vB$riverOrdered, sep = '_'), ordered = T)

@@

vB$salmon01 <- ifelse(vB$cohort <= 2004 & vB$river == 'west brook', 1, 0)

vB$speciesGG <- factor(vB$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
vB$seasonGG <- factor(vB$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)

# scaling temperature and flow
vB$meanTemperatureScaled <- as.numeric(scale(vB$meanTemperature))
vB$meanFlowScaled <- as.numeric(scale(vB$meanFlow))
vB$meanTemperatureFlowScaled <- vB$meanTemperatureScaled * vB$meanFlowScaled

seasonalTF <- vB %>% group_by(season) %>%
  summarize(meanTBySeason = mean(meanTemperature, na.rm = TRUE),
            sdTBySeason = sd(meanTemperature, na.rm = TRUE),
            meanFBySeason = mean(meanFlow, na.rm = TRUE),
            sdFBySeason = sd(meanFlow, na.rm = TRUE))

vB <- left_join(vB,seasonalTF) %>%
  mutate(meanTemperatureScaledBySeason = (meanTemperature - meanTBySeason)/sdTBySeason, 
         meanFlowScaledBySeason = (meanFlow - meanFBySeason)/sdFBySeason)

vB$meanTemperatureFlowScaledBySeason <- vB$meanTemperatureScaledBySeason * vB$meanFlowScaledBySeason
```

@@

ggplot(cohortStrength, aes(cohort, cohortStrengthScaled, group = species, linetype = species, shape = species, color = species)) +
  geom_point(size = 2) +
  geom_line(size = 1.1, color = 'grey40') +
  labs(x = 'Cohort',
       y = 'Cohort strength') +
  theme_publication() +
  facet_grid(~metaPop01)

@@

```{r raw data histos}

ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt'), river == "west brook"), aes(observedLength)) +
  geom_freqpoly(aes(color = species)) +
  theme_publication() +
  facet_grid(season ~ year)

# season 3
ggplot(vB %>% filter(cohort %in% 2000:2013, species %in% c('bkt', 'bnt'), river == "west brook", season == 3), aes(observedLength)) +
  geom_freqpoly(aes(color = speciesGG), binwidth = 6, size = 1.35) +
  theme_publication() +
  theme(strip.text.x = element_text(size = 8), 
        strip.background = element_blank(),
        legend.title = element_text("Species"),
        legend.position = "right",
        legend.direction = "vertical") + 
  labs(color='Species') +
  scale_x_continuous("Fork Length (mm)", limits = c(49,200)) +
  scale_y_continuous("Count") +
  #scale_color_manual(values = c("lightblue", "lightgreen")) +
  scale_colour_brewer(palette = "Paired")+
  facet_wrap(~ year)

ggsave('./figures/season3freqpoly.png', width = 11, height = 8, units='in')

# 1 cohort
ggplot(vB %>% filter(cohort == 2009, species %in% c('bkt', 'bnt'), river == "west brook"), aes(observedLength)) +
  geom_freqpoly(aes(color = speciesGG), binwidth = 6, size = 1.35) +
  theme_publication() +
  theme(strip.text.x = element_text(size = 8), 
        strip.background = element_blank(),
        legend.title = element_text("Species"),
        legend.position = "right",
        legend.direction = "vertical") + 
  labs(color='Species') +
  scale_x_continuous("Fork Length (mm)", limits = c(49,200)) +
  scale_y_continuous("Count") +
  #scale_color_manual(values = c("lightblue", "lightgreen")) +
  scale_colour_brewer(palette = "Paired")+
  facet_grid(season ~ year)

ggsave('./figures/cohort2009freqpoly.png', width = 11, height = 8, units='in')

```{r mean lengths: from MS_Analysis_3spp_WB.Rmd. For Grenada talk}

# Has mean size changed over time?
means <- 
  cd %>% mutate( age = year - cohort ) %>%
    filter( riverOrdered == 'west brook', ageInSamples %in% 1:10, age < 3, year < 2016 ) %>%
    group_by(species,ageInSamples,year,season,age) %>%
    summarize(meanLen = mean(observedLength,na.rm = TRUE),
              sdLen = sd(observedLength,na.rm = TRUE),
              meanGR = mean(grLength,na.rm = TRUE),
              sdGR = sd(grLength,na.rm = TRUE)
              )

lenSlopes <- means %>%
  group_by( species, ageInSamples, season, age ) %>%
  nest()

lengthModel <- function(df){
  lm( meanLen ~ year, data = df )
}

grModel <- function(df){
  lm( meanGR ~ year, data = df )
}

lenSlopes <- lenSlopes %>%
  mutate( model = map(lenSlopes$data, lengthModel),
          modelGR = map(lenSlopes$data, grModel))

###################################
######### len here, gr below ######
lenSlopesTidy <- lenSlopes %>%
  mutate(glance = map(model, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig))

means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
lenSlopesTidy$seasonGG <- factor(lenSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
lenSlopesTidy$speciesGG <- factor(lenSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanLen, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean body length (mm)" ) +

  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age<2), 
            aes(x=2000,y=225, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age<2), 
            aes(x=2000,y=195, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +

  geom_text(data = lenSlopesTidy %>% filter(species == 'bkt', age==2), 
            aes(x=2011.8,y=104, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
  geom_text(data = lenSlopesTidy %>% filter(species == 'bnt', age==2), 
            aes(x=2011.8,y=74, label=paste(round(estimate,2),sig)),hjust = 0, size = 6)+

  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
      #  legend.position = "right",
     #   legend.direction = "vertical"
        legend.position = "bottom",
        legend.direction = "horizontal") + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age, scales='free')

ggsave('./figures/meanSizes.png', width = 11, height = 11, units='in')

# one season/year only
ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 8), aes(year,meanLen, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean body length (mm)" ) +
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
      #  legend.position = "right",
     #   legend.direction = "vertical"
                legend.position = "bottom",
        legend.direction = "horizontal") + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age)
ggsave('./figures/meanSizesAIS_1.png', width = 11, height = 8, units='in')

# is a model with salmon presence absence better than a year model?
means$ats01 <- ifelse(means$year <= 2005, 1, 0)

modMeans1 = lm(meanLen ~ year * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
modMeans2 = lm(meanLen ~ year * factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
modMeans3 = lm(meanLen ~ factor(ats01) * species, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
modMeans4 = lm(meanLen ~ factor(ats01)*year, data = means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1))
AIC(modMeans1, modMeans2, modMeans3, modMeans4)

# one season/year only
ggplot( means %>% filter(year >= 2000, species != 'ats', ageInSamples == 1), aes(year,meanLen, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(group = factor(ats01)), color = 'black', se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean body length (mm)" ) +
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
      #  legend.position = "right",
     #   legend.direction = "vertical"
        legend.position = "bottom",
        legend.direction = "horizontal") + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age)

ggsave('./figures/meanSizesAIS_1_ats01.png', width = 11, height = 8, units='in')

# mean increases 
lenSlopesTidy %>%
  filter(species != 'ats') %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
lenSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
ggplot(lenSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point(size = 5) +
  geom_smooth(method='lm', se=FALSE, size = 1.5) +
  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
  labs( x = "Age Season", y = "Slope of mean body length over years" ) +
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
        legend.position = "bottom",
        legend.direction = "horizontal",
        axis.text.x = element_text(angle = 40, hjust = 1)) + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75)

ggsave(paste0('figures/meanLengthsOverAge.png'), width = 11, height = 10, units='in')

# species diff in increases in length?
sppEffect <- (lm(estimate ~ ageInSamples * (species), data = lenSlopesTidy))
summary(sppEffect)
anova(sppEffect)

# mean diff in slopes btw bkt and bnt
lenSlopesTidy %>%
  filter(species != 'ats') %>%
  group_by(species) %>%
  summarise(means = mean(estimate))

#####################################  
#################################
# growth rate
grSlopesTidy <- lenSlopes %>%
  mutate(glance = map(modelGR, broom::tidy)) %>% 
  unnest(glance, .drop = TRUE) %>%
  filter(term == 'year') %>%
  mutate( sig = ifelse(p.value < 0.05,"*","")) %>%
  mutate( sig = ifelse(p.value < 0.001,"**",sig)) %>%
  mutate( sig = ifelse(p.value < 0.0001,"***",sig))

##
means$seasonGG <- factor(means$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
means$speciesGG <- factor(means$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)
grSlopesTidy$seasonGG <- factor(grSlopesTidy$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
grSlopesTidy$speciesGG <- factor(grSlopesTidy$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot( means %>% filter(year >= 2000, species != 'ats'), aes(year,meanGR, group= speciesGG)) +
  geom_point( aes(color = speciesGG), size = 4) +
  geom_smooth( aes(color = speciesGG), se = FALSE, method = 'lm') +
  labs( x = "Year", y = "Mean growth rate (mm/d)" ) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age<2), 
            aes(x=2000,y=0.68, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age<2), 
            aes(x=2000,y=0.55, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +

  geom_text(data = grSlopesTidy %>% filter(species == 'bkt', age==2), 
            aes(x=2011.5,y=0.64, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+#, color = speciesGG)) +
  geom_text(data = grSlopesTidy %>% filter(species == 'bnt', age==2), 
            aes(x=2011.5,y=0.51, label=paste(round(estimate,4),sig)),hjust = 0, size = 6)+
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
      #  legend.position = "right",
     #   legend.direction = "vertical"
        legend.position = "bottom",
        legend.direction = "horizontal") + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG~age)

ggsave('./figures/meanGRs.png', width = 11, height = 11, units='in')

# mean increases 
grSlopesTidy %>%
  filter(species != 'ats') %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# mean increases by species
grSlopesTidy %>%
  group_by(species) %>%
  summarise( means = mean(estimate),
             sd = sd(estimate))

# do increases (slopes) increase over age? Yes
ggplot(grSlopesTidy %>% filter(species != 'ats'), aes((ageInSamples),estimate, color=speciesGG)) +
  geom_point(size = 5) +
  geom_smooth(method='lm', se=FALSE) +  
  geom_hline(yintercept = 0) +
  scale_x_continuous(breaks = 1:10, labels = c('0 Fall','0 Winter','1 Spring','1 Summer','1 Fall','1 Winter','2 Spring','2 Summer','2 Fall','2 Winter')) +
  labs( x = "Age Season", y = "Slope of mean growth rate over years" ) +
  theme_publication(base_size = 24) +
  theme(legend.title = element_text("Species"),
        legend.position = "bottom",
        legend.direction = "horizontal",
        axis.text.x = element_text(angle = 40, hjust = 1)) + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75)

ggsave('./figures/meanGRsOverAge.png', width = 11, height = 11, units='in')

summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bkt')))
summary(lm(estimate~ageInSamples, data = grSlopesTidy %>% filter(species=='bnt')))

# species diff in increases in gr?
sppEffectGR <- (lm(estimate ~ ageInSamples * (species), data = grSlopesTidy))
summary(sppEffectGR)
anova(sppEffectGR)

# mean diff in slopes btw bkt and bnt
grSlopesTidy %>%
  filter(species != 'ats') %>%
  group_by(species) %>%
  summarise(means = mean(estimate))

```{r size-dep imm or mort? For Spain talk}

### for losses

cd01 <-  cd %>%
  filter(knownZ == 1) %>%
  group_by(tagIndex) %>%
  mutate(first = min(sampleNumber),
         firstObs = sampleNumber == first,
         last = max(sampleNumber),
         lastObs = sampleNumber == last,
         notLastObs = !lastObs,
         nObs = n(),
         age = year - cohort
  ) %>%
  ungroup()

meansByAIS <- cd01 %>%
  group_by(species, ageInSamples, cohort, river) %>%
  summarise(meanLen = mean(observedLength, na.rm =T),
            sdLen = sd(observedLength, na.rm = T),
            n = n())
cd01 <- left_join(cd01, meansByAIS) %>%
  mutate(observedLengthStd = (observedLength - meanLen) / sdLen,
         cohortStd = cohort - 2007)

#cd %>% select(tagIndex,sampleNumber, firstObs,lastObs, knownZ, observedLength) %>% head(50) %>% as.data.frame()
cd01Mod <- cd01 %>% filter(species %in% c('bkt', 'bnt'), river == 'west brook', ageInSamples %in% 1:10, cohort %in% 2000:2014)

ggplot(cd01Mod, aes(observedLengthStd, notLastObs*1, color = factor(species), group = factor(species))) +
  geom_point(alpha = 0.01) +
  geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE) +
 # scale_color_viridis_c() +
  facet_wrap( ~ cohort, scales = 'free')

#### by species and ageInSamples

sensSlopes <- cd01Mod %>%
  group_by( species, ageInSamples ) %>%
  nest()

survModel <- function(df){
  glm(notLastObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
}
enterModel <- function(df){
  glm(firstObs*1 ~ observedLengthStd + cohortStd, family = 'binomial', data = df )
}

sensSlopes <- sensSlopes %>%
  mutate(modelSurv = map(sensSlopes$data, survModel),
         modelEnter = map(sensSlopes$data, enterModel))

sensSlopesTidy <- sensSlopes %>%
  mutate(surv = map(modelSurv, broom::tidy),
         enter = map(modelEnter, broom::tidy)) %>% 
  unnest(surv, enter, .drop = TRUE) %>%
  filter(term == 'cohortStd') %>% # cohort effect
  rename(estimateSurv = estimate, estimateEnter = estimate1) %>%
  select(species,ageInSamples,estimateSurv, estimateEnter) #%>%
  #gather("type", "estimate", estimateSurv:estimateEnter)

ggplot(sensSlopesTidy, aes(ageInSamples, estimate, color = type)) +
  geom_point() +
  geom_line() +
  facet_wrap(~species)

```{r dummy graphs for Spain talk}

sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
ggplot(sizesForGraph, aes(year,surv)) +
  geom_line(size = 2) +
  labs(x = 'Year', y = 'Mean body length (mm)') +
  ylim(80,200) +
  theme_publication(base_size = 36)
ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')

survForGraph <- data.frame(size = c(-2,2), surv = c(0.25,0.75))
ggplot(survForGraph, aes(size,surv)) +
  geom_line(size = 2) +
  labs(x = 'Standardized length', y = 'Probability of seen again') +
  theme_publication(base_size = 36)
ggsave('./figures/pSurvivalDummy.png', width = 11, height = 11, units='in')

firstForGraph <- data.frame(size = c(-2,2), surv = c(0.35,0.65))
ggplot(firstForGraph, aes(size,surv)) +
  geom_line(size = 2) +
  labs(x = 'Standardized length', y = 'Probability of first observation') +
  theme_publication(base_size = 36)
ggsave('./figures/pFirstDummy.png', width = 11, height = 11, units='in')

sizesForGraph <- data.frame(year = c(2000,2015), surv = c(100, 150))
ggplot(sizesForGraph, aes(year,surv)) +
  geom_line(size = 3) +
  labs(x = 'Year', y = 'Mean body length (mm)') +
  ylim(80,200) +
  theme_publication(base_size = 36)
ggsave('./figures/sizesDummy.png', width = 11, height = 11, units='in')

#### show interactions

survForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.52,0.55,0.51,0.24,0.54,0.82))
ggplot(survForGraph, aes(size,surv, group = ind)) +
  geom_line(size = 2) +
  labs(x = 'Standardized length', y = 'Probability of seen again') +
  theme_publication(base_size = 36)
ggsave('./figures/pSurvivalDummy3.png', width = 11, height = 11, units='in')

firstForGraph <- data.frame(ind = c(1,1,2,2,3,3), size = c(-2,2), surv = c(0.42,0.45,0.41,0.14,0.44,0.72))
ggplot(firstForGraph, aes(size,surv, group = ind)) +
  geom_line(size = 2) +
  labs(x = 'Standardized length', y = 'Probability of first observation') +
  theme_publication(base_size = 36)
ggsave('./figures/pFirstDummy3.png', width = 11, height = 11, units='in')

```{r get betas from growth model for sensitivities}
load("D:/projects/linkedModels/data/out/allBetas.RData") #loads df 'q'
qFig1 <- q %>% filter(riverGG == 'WB', species %in% c('bkt','bnt'))
qFig1$season <- as.numeric(as.character(qFig1$season))
qFig1$isYOY <- as.numeric(as.character(qFig1$isYOY))
qFig <- qFig1 %>%
  mutate( varText = tolower(betaGG)) %>%
  mutate( varText = ifelse(varText == 'temp*flow','temp:flow', varText)) %>%
  mutate( varText = ifelse(varText == 'bnt*bkt','bkt:bnt', varText)) %>%
  select(isYOY,season,species,median,varText)
qFig$seasonGG <- factor(qFig$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
#get slopes of env variables

load('D:/projects/linkedModels/data/out/flowSlopesTidy.RData')
flowSlopesTidy2 <- flowSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'flow')
load('D:/projects/linkedModels/data/out/tempSlopesTidy.RData')
tempSlopesTidy2 <- tempSlopesTidy %>% filter(riverOrderedGG == 'West Brook') %>% select(estimate, seasonGG) %>% mutate(varText = 'temp')
load('D:/projects/linkedModels/data/out/nSlopesTidy.RData')
nSlopesTidy2 <- nSlopesTidy %>% select(estimate, season, species) %>% mutate(varText = species)
nSlopesTidy2$seasonGG <- factor(nSlopesTidy2$season, levels = 1:4, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
nSlopesTidy2 <- nSlopesTidy2 %>% select(estimate,seasonGG,varText)

allSlopesTidy <- bind_rows(nSlopesTidy2, tempSlopesTidy2, flowSlopesTidy2)

allSlopesTidyWBetas <- left_join(qFig, allSlopesTidy, by = c('seasonGG', 'varText')) %>% mutate(sens = median * estimate)

ggplot(allSlopesTidyWBetas%>%filter(!is.na(sens), varText != 'ats'), aes(seasonGG, sens, color = varText)) +
  geom_point() +
  facet_grid(species + isYOY ~ varText, scales = 'free')
```

```{r }
############# 1st attempt at imm em
mod1 <- glm(notLastObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
#mod2 <- glm(notLastObs*1 ~ observedLengthStd * factor(cohortStd) * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
mod3 <- glm(notLastObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)

AIC(mod1,mod3)

summary(mod1)
anova(mod1, test="Chi")

pred <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))

pred$pred <- predict.glm(mod1, newdata = pred, type = 'response')

u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
pred <- left_join(pred, u)

pred$seasonGG <- factor(pred$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
pred$cohort <- pred$cohortStd + 2007
pred$speciesGG <- factor(pred$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot(pred, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
  geom_point() +
  scale_color_viridis_c() +
  scale_y_continuous("Probability of survival") +
  facet_grid(species ~ ageInSamples, scales = 'free')

ggplot(pred %>% filter(cohortStd %in% 0, age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
  geom_point(size = 2) +
  scale_x_continuous("Standardized length") +
  scale_y_continuous(lim = c(0,1),"Probability of seen again") +
  theme_publication(base_size = 28) +
  theme(legend.title = element_text("Species"),
        legend.position = "bottom",
        legend.direction = "horizontal"
        ) + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age)

ggsave('./figures/pSurvival.png', width = 11, height = 11, units='in')

################################
# gains - immigration, firstObs

ggplot(cd01Mod, aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE) +
 # scale_color_viridis_c() +
  xlim(c(-2.75,2.75)) +
  facet_grid(ageInSamples ~ cohort)

ggplot(cd01Mod%>%filter(ageInSamples == 3), aes(observedLengthStd, firstObs*1, color = factor(species), group = factor(species))) +
  geom_jitter(alpha = 0.1) +
  geom_smooth(method = "glm", 
    method.args = list(family = "binomial"), 
    se = FALSE) +
 # scale_color_viridis_c() +
  xlim(c(-2.75,2.75)) +
  facet_wrap( ~ cohort)

modi1 <- glm(firstObs*1 ~ observedLengthStd * cohortStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
modi2 <- glm(firstObs*1 ~ observedLengthStd * species * factor(ageInSamples), family = 'binomial', data = cd01Mod)
modi3 <- glm(firstObs*1 ~ observedLengthStd * species , family = 'binomial', data = cd01Mod)
modi4 <- glm(firstObs*1 ~ observedLengthStd * factor(ageInSamples), family = 'binomial', data = cd01Mod)
modi5 <- glm(firstObs*1 ~ observedLengthStd, family = 'binomial', data = cd01Mod)

AIC(modi1,modi2,modi3,modi4,modi5) %>% rownames_to_column() %>% arrange(AIC)

summary(modi1)
anova(modi1, test="Chi")
###
predi <- expand.grid(observedLengthStd = seq(-2,2, length.out = 50), cohortStd = unique(cd01Mod$cohortStd), species = c('bkt', 'bnt'), ageInSamples = unique(cd01Mod$ageInSamples))

predi$pred <- predict.glm(modi1, newdata = predi, type = 'response')

u <- cd01Mod %>% distinct(age, season, ageInSamples) %>% arrange(ageInSamples) %>% filter(!(age==2 & ageInSamples == 6))
predi <- left_join(predi, u)

ggplot(predi, aes(observedLengthStd, pred, color = (cohortStd), group = (cohortStd))) +
  geom_point() +
  scale_color_viridis_c() +
  scale_y_continuous("Probability of first observation") +
  facet_grid(species ~ ageInSamples, scales = 'free')

predi$seasonGG <- factor(predi$season, labels = c("Spring","Summer","Autumn","Winter"), ordered = T)
predi$cohort <- predi$cohortStd + 2007
predi$speciesGG <- factor(predi$species, levels = c('bkt','bnt','ats'), labels = c("Brook Trout", "Brown Trout", "Atlantic Salmon"), ordered = T)

ggplot(predi %>% filter(cohortStd %in% c(0), species == 'bkt', age < 3), aes(observedLengthStd, pred)) +
  geom_point() +
  scale_color_viridis_c() +
  scale_x_continuous("Standardized length") +
  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
  theme_publication() +
  facet_grid(seasonGG ~ age)

ggplot(predi %>% filter(cohortStd %in% c(0), age < 3), aes(observedLengthStd, pred, color = speciesGG)) +
  geom_point(size = 2) +
  scale_x_continuous("Standardized length") +
  scale_y_continuous(lim = c(0,1),"Probability of first observation") +
  theme_publication(base_size = 28) +
  theme(legend.title = element_text("Species"),
        legend.position = "bottom",
        legend.direction = "horizontal"
        ) + 
  labs(color='Species') +
  scale_color_grey(start = 0.15, end = 0.75) +
  facet_grid(seasonGG ~ age)

ggsave('./figures/pFirstObs.png', width = 11, height = 11, units='in')
```

```{r gam models - find best lm, before fitting detailed gams - only need to run for model selection}
# only including river, season, salmon01 becaseu we assume these have categorical effects

mod0 <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod1 <- bam(observedLength ~ riverFactor + factor(season) + factor(salmon01) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod2 <- bam(observedLength ~ riverFactor * factor(season) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod3 <- bam(observedLength ~ riverFactor + factor(season) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod4 <- bam(observedLength ~ riverFactor + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod5 <- bam(observedLength ~ factor(season) + 
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

mod6 <- bam(observedLength ~ riverFactor + factor(salmon01) +
               s(ydayCumul, bs = 'cr'), 
               data = dataIn, 
               method = "REML")

AIC(mod0,mod1,mod2,mod3,mod4,mod5,mod6) %>% rownames_to_column() %>% arrange(AIC)
summary(mod0)
plot(mod0)
gam.check(mod0)

```{r gam structure given best model from above}

# BKT, mod0 is best fixed effects model

system.time(
mod0_YDCcr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor), 
               data = dataIn, 
               method = "REML")
)

system.time(
mod0_YDCc_r <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) , 
               data = dataIn, 
               method = "REML")
)

AIC(mod0_YDCcr, mod0_YDCc_r) %>% rownames_to_column() %>% arrange(AIC)

system.time(
mod0_YDCcr_Cr_CSr_Tr_Fr_TFr <- bam(observedLength ~ riverFactor * factor(season) + factor(salmon01) +              
               s(ydayCumul, bs = 'cr') +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = riverFactor) +
               s(ydayCumul, bs = 'cr', k = kIn, by = cohortRiverFactor) +
               s(cohort, bs = 'cr', k = kIn, by = riverFactor) + 
               s(cohortStrengthScaled, bs = 'cr', k = kIn, by = riverFactor) +  # this is conspecific cS only
               s(meanTemperatureScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor) +
               s(meanTemperatureFlowScaledBySeason, bs = 'cr', k = kIn, by = riverFactor), 
               data = dataIn, 
               method = "REML")
)
